Setting up Data

### 2020-02-04 doc/Main_Report_Results.rmd
summarizing all results, updating saving files

# Results

@@

Details of data set cleaning etc is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd

@@

# A. Models for av MLH1 counts foci per cell

ToDo: change the Mixed models to lmer() so that I can confidantly test random effect (variance)

@@

$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)  + \varepsilon$$

I won't run the model above -- even tho it is logical and makes sense -- I can't make good estimates for random effects.

$$mouse .var CO ~=~ subsp * sex + rand(strain) + \varepsilon $$

@@

- Unorder factors

- male ages within good range (5 - 12? weeks)

- Mol as a different subsp

(Move this into the Methods)

*The mouse average table used in this part excluded Q5 cells*

*the MLH1 mean is from mean - adj_MLH1 (males have +1 CO to count for PAR)*

@@

subspecies levels: 2(3), Musc and Domestics and Molossinus

strain levels: Don: 3, Musc: 4  Mol: 2

@@

*mouse averages exclude Q5 and male counts are adjusted +1 for PAR)*

## Mean CO Count

todo, add the random effect tests, consider changing all to lmer() (those have better random effects)

For now I will focus on 4 different models run mouse averages on MLH1 counts.

1. $$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$

USE the bottom one

2. $$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

3. $$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$

4. $$mouse .mean CO ~=~ sex * strain + \varepsilon$$

> Model 1

$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$

1.	HetC.MixedModel.HQ <- lme(mean_co ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )

```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2, eval=FALSE}
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
#mean_co.HQ  (this might remove some of the MOLF males)

HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )

#ranef(HetC.MixedModel.HQ)
#which is the test for random effects ?!
```

Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.  
-->

> Model 2. THIS SHOULD BE THE MAIN MIXED MODEL --lmer

@@

```{r MixedModel.1, echo=FALSE, digits = 3}
#THIS IS THE NEW MIXED MODEL
#should be able to use the loaded 
#DF.HetC.MixedModel.HQ

@@

summary(MLH1.MM.M1)
drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***
#only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
anova(MLH1.MM.M1)

@@

as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
#add a aption to this code chunk
#ranef(Reduced.strain.HQ)
```

@@

Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
This is model has sex and interaction coefficient is significant.

Can the above model be re-written with lmer so that random effect can be tested?

> Model 3

All fixed effects with all interactions.

$$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$

```{r MixedModel.3, echo=FALSE}

#strain = fixed
lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
#library(xtable)

@@

> Model 4

$$mouse .mean CO ~=~ sex * strain + \varepsilon$$

```{r MixedModel.4, echo=FALSE}

@@

Above is the table of coefficients for **MM.4**, HQ data set, lm function, just sex and strain, no subspecies.

Significant factors are 
strain G (p=0.0005), strainMSM and strainPWD (very slight),
sex:G (p=0.001),
sex:MSM, Sex:PWD (very slight p=.05, p=0.3),

#### ANOVA, mean CO count full vs reduced

@@

Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
Models with random effects more significant than nulls (both just strain and no random effects). The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.

```{r Base.strain.fixed, echo=FALSE}
#make mixed model with strain as fixed, since there are interesting results
#Consider including these results from all FIXED MODEL

HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
#this produces several significant factors

HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )

#summary(HetC.MM.strain.fixed2)
#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))

#summary(HetC.MM.strain.fixed)
#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **

#male effect lowers the mean from intercept
#while strain effects 
```

 Summary of mean COs

MM.2 has significant coefficients, sex and sex*subsp interaction.

MM.2.Mol, the sex and interaction term (subsp*sex) are significant.

MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.

## Musc Males

```{r Musc.male.Model.data.setup, echo=FALSE}
#nested mouse in strain

#remove the extra merged cols
Musc.male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

Musc.male_mouse.av <- add_strain(Musc.male_mouse.av)
Musc.male_mouse.av <- add_sex(Musc.male_mouse.av)
Musc.male_mouse.av <- add_subsp(Musc.male_mouse.av)
Musc.male_mouse.av <- add_species(Musc.male_mouse.av)

Musc.male_mouse.av$subsp <- as.character(Musc.male_mouse.av$subsp)

Musc.male_mouse.av$subsp[Musc.male_mouse.av$strain== "MSM"] <- 'Mol'
Musc.male_mouse.av$subsp[Musc.male_mouse.av$strain== "MOLF"] <- "Mol"

#use original for getting the other strains

Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$species == "M.musculus",]
Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$subsp == "Musc",]
Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$sex == "male",]

#set the factors
Musc.male_mouse.av$strain <- factor(Musc.male_mouse.av$strain, ordered = FALSE)
Musc.male_mouse.av$mouse <- factor(Musc.male_mouse.av$mouse, ordered = FALSE)

#compare the the higher quality plots
#Q1Q2

#Q12_mouse_table_w.Ages

Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_species(Q12_mouse_table_w.Ages)

Q12_mouse_table_w.Ages$subsp <- as.character(Q12_mouse_table_w.Ages$subsp)

Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MSM"] <- 'Mol'
Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MOLF"] <- "Mol"

Musc.male.Q12 <- Q12_mouse_table_w.Ages[(Q12_mouse_table_w.Ages$subsp == "Musc")&(Q12_mouse_table_w.Ages$sex == "male"),]

Musc.male.Q12$strain <- factor(Musc.male.Q12$strain, ordered = FALSE)
Musc.male.Q12$mouse <- factor(Musc.male.Q12$mouse, ordered = FALSE)

#Q123  Q123_mouse_table

Q123_mouse_table <- add_subsp(Q123_mouse_table)
Q123_mouse_table <- add_species(Q123_mouse_table)

Q123_mouse_table$subsp <- as.character(Q123_mouse_table$subsp)

Q123_mouse_table$subsp[Q123_mouse_table$strain== "MSM"] <- 'Mol'
Q123_mouse_table$subsp[Q123_mouse_table$strain== "MOLF"] <- "Mol"

Musc.male.Q123 <- Q123_mouse_table[(Q123_mouse_table$subsp == "Musc")&(Q123_mouse_table$sex == "male"),]

Musc.male.Q123$strain <- factor(Musc.male.Q123$strain, ordered = FALSE)
Musc.male.Q123$mouse <- factor(Musc.male.Q123$mouse, ordered = FALSE)

```{r Musc.male.show.plots, echo=FALSE}

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)

# + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +

#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +  
musc.male.plot

Q123.musc.male.plot <- ggplot(Musc.male.Q123,
                             aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q123 Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Q123.musc.male.plot
#THE q12,data set only 
#Q123 has 5 strain rep (no TOM), there are still some mouse outliers

```{r M1_Musc.male, echo=FALSE}

#mouse.av ~ mouse | strain

#not sure the right syntax for nested factors
#Musc.male.M1 <- lm(mean_co ~ (mouse|strain), data=Musc.male_mouse.av )

Musc.male.M1 <- lm(mean_co ~ strain, data=Musc.male_mouse.av )

as.data.frame(  round(summary(Musc.male.M1)$coefficients, 4) )

#all of the strain have a significant effect agaisnt the group mean / intercept
#looking at the intercept -- the PWD observations are pushing up the average

#musc male Q123  Q123_mouse_table

Musc.male.Q123.M1 <- lm(mean_co ~ strain, data=Musc.male.Q123 )

as.data.frame(  round(summary(Musc.male.Q123.M1)$coefficients, 4) )

#mostly all strains, (but SKIVE) are significant -- similar to the top model

#shoul I compare with subsampling

```{r Musc.male.M2}

#mouse.av ~ strain

@@

>>>>

@@

<!--  insert Details of data set cleaning etc is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd -->

@@

# Models Framework

@@

$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$

@@

- Male gwRR have been adjusted for the PAR

- Q5 excluded

- All the effects are unorder factors

- male ages within 5 - 20 weeks

- Molossinus is coded as a subsp level

@@

subspecies levels: 3, Musc, Dom and Molossinus

strain levels: Dom: 3, Musc: 3  Mol: 2

@@

For now I will focus on 3 different models for comparing the subsp, sex and strain effects across relevant dependant variables.

1. $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

Allows random effect from strain. Considers the wild derived inbred strains as random samples from each subspecies territory.

2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$

Post hoc model for investigating strain as a fixed effect.

3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$

Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.

## MLH1 count M1

@@

```{r MLH1co.av.M1, echo=FALSE, digits = 3}
#THIS IS THE NEW MIXED MODEL
#should be able to use the loaded 
#DF.HetC.MixedModel.HQ

@@

#summary(MLH1.MM.M1)
drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***

as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))

#only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
anova(MLH1.MM.M1)

@@

@@

## MLH1 count M2

$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$

```{r MixedModel.2, echo=FALSE,  digits = 3}
#strain = fixed
lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)

anova(lm.strain.fixed.full)

@@

## MLH1 count M3

$$mouse \ av \ metric ~=~ sex * strain + \varepsilon$$

```{r MixedModel.3, echo=FALSE, digits = 3}

@@

<!--
 Contrasts and ANOVA, mean CO count full vs reduced

@@

@@

## Male Specific Polymorphism

Follow up on this finding -- included more strains which were excluded from the HetC analysis.

-- what's the driving question for this section --

> is there a strain effect in Dom?

> How different are the Musc strains?

```{r male.show.plots, echo=FALSE}

male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)

fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)

table( droplevels(fomo.male_mouse.av$strain) )
#there are cast strains too

Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
musc.male.plot

# + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +
#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +

#make the Mol and
Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")

Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
Mol.male.plot

#Dom version 
Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")

Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
Dom.male.plot

```{r Male.specific.test, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect

#1. determine which strain effects are sig

bhh11 <- lm(mean_co ~ subsp * strain, data= fomo.male_mouse.av)
#bhh11$coefficients

booo <- summary(bhh11)
booo$coefficients

lm.male.strain <- lm(mean_co ~ strain, data= fomo.male_mouse.av)
ll <- summary(lm.male.strain)
ll$coefficients
#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

msm, pwd skive are when all other male strains are added. skive is less sig than the others -- but it's still sig.  Given these results I can group SKIVE, PWD and MSM into the 'high rec' group.

*PWD, MSM, SKIVE are in 'high rec group'*

### 2020-02-04 src/SetupData.Rmd
summarizing all results, updating saving files

@@

@@

```{r exclude.mice, echo=FALSE}
#remove mice -- 18nov17_WSB_f5 (only 1 cell)
exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
                  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                  "3sep18_AST_m1",
                  "3jul18_TOM_m1","3sep18_TOM_m1",
                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]

@@

@@

                             Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)

)

@@

```{r Q123_mouse.table, echo=FALSE}

@@

>>>>

-rm Peromycus data BivData!!

@@

- rm all of BD data?

@@

```{r exclude.mice, echo=FALSE}
#remove mice -- 18nov17_WSB_f5 (only 1 cell)
exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                 "3sep18_AST_m1",
                  "3sep18_TOM_m1",
                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]

@@

the excluded TOM, had low quality cells, only 4. The exclude AST, had low poor quality cells, there are 3 other mice.  The excluded CZECH males are poor quality (but near the mean). There are 3 other males.

@@

                             Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

@@

```{r Q123_mouse.table, echo=FALSE, eval=FALSE}

@@

```{r male.specific.table,echo=FALSE}

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels

male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")

#formate the table for pluggin into models
